Spatial and temporal variability in economic availability of residual biomass

### 2021-12-20 analysis/newms_elsevier.Rmd
include word docs

Humans have been using biofuels since we learned to control fire a
million or so years ago.  In developed economies, biofuels were
largely displaced by coal during the industrial revolution, and by
petroleum starting in the mid-nineteenth century.  In recent times,
biofuels are an increasingly important source of energy because of
their potential to reduce dependency on fossil fuels and reduce
greenhouse gas emissions.  Direct combustion of biomass is still an
important source of energy.  In Europe, in particular, many
electricity plants have been converted to use wood pellets instead of
coal as the primary energy source.  Liquid biofuels (*e.g.* ethanol
and biodiesel) have been produced in commercial quantities from food
crops starting in the 1980s.  These are called first generation
biofuels.  A major difficulty with these first generation biofuels is
that they compete with food and that their use  may lead to land use change.

@@

The ATS boundaries are particularly important to this study as all the
data used for this study are aggregated or disaggregated to the
township level.  ATS provides the basis for legal land descriptions in
Alberta.  The typical township is roughly square (9.778 km by 9.716
km, 9 500 ha) and is indexed by township (1--126), range (1--30), and
meridian (4--6).  The townships immediately east of Alberta's western
boundary, and immediately east of the 5^th^ and 6^th^ meridians
(corresponding to 114°W and 118°W longitude, respectively) are partial
townships and therefore deviate from the roughly square shape.  ATS is
described in detail by the Alberta Land Surveyor's Association
@ALSA2021.  There are 7 237 townships in Alberta, allowing for spatial
resolution appropriate for regional or provincial level analyses.

@@

This is a greedy algorithm.  It ignores the cost of moving equipment
from one block to another, but we believe it still provides useful
information.

@@

All costs and prices used in this study are in Canadian dollars (\$).
At the time of writing (`r sprintf("%s",fx[1,1])` GMT), the exchange rate between
Canadian (CAD) and US dollars (USD) was `r sprintf("%.4f",fx[1,2])` USD/CAD.

@@

The bulk of the effort expended on this study was spent in creating
township-level summaries for use by the model.  The model itself is
very simple.  We assume that the residue harvesters follow a kind of
greedy optimization.  They take the cheapest residue first, and stop
when they have enough residue to satisfy annual plant
requirements. For each combination of plant location and year, all the
townships in Alberta were sorted by ascending cycle time (equivalently
tranport cost).  The model works through this list, accumulating harvest
residue recovered from the harvest area associated for the township
and year, and provided a file showing cycle time, total cost, and
cumulative residue recoverd.  Once this file is created, fairly simple
data manipulations are undertaken to produce the results.

@@

@@

The data compiled to the township level are currently available on
github(https://github.com/gwa-uab/histressup/tree/main/data).  Before
publication it will be placed on ERA with a DOI.

>>>>

Humans have been using biofuels since we learned to control fire a
million or so years ago.  In developed economies, biofuels were
largely displaced by coal during the industrial revolution, and by
petroleum starting in the mid-nineteenth century.  In recent times,
biofuels are an increasingly important source of energy because of
their potential to reduce dependency on fossil fuels and reduce
greenhouse gas emissions.  Direct combustion of biomass is still an
important source of energy.  In Europe, in particular, many
electricity plants have been converted to use wood pellets instead of
coal as the primary energy source.  Liquid biofuels (*e.g.* ethanol
and biodiesel) have been produced in commercial quantities from food
crops starting in the 1980s.  These are called first generation
biofuels.  A major difficulty with these first generation biofuels is
that they compete with food and that their use may lead to increased
food costs land use change.

@@

The ATS boundaries are particularly important to this study as all the
data used for this study are aggregated or disaggregated to the
township-level.  ATS provides the basis for legal land descriptions in
Alberta.  The typical township is roughly square (9.778 km by 9.716
km, 9 500 ha) and is indexed by township (1--126), range (1--30), and
meridian (4--6).  The townships immediately east of Alberta's western
boundary, and immediately east of the 5^th^ and 6^th^ meridians
(corresponding to 114°W and 118°W longitude, respectively) are partial
townships and therefore deviate from the roughly square shape.  ATS is
described in detail by the Alberta Land Surveyor's Association
@ALSA2021.  There are 7 237 townships in Alberta, allowing for spatial
resolution appropriate for regional or provincial level analyses.

@@

The goal of the data aggregation is to provide a data set for a greedy 
optimatizion model that minimizes the cost of harvesting forest residues
for combinations of plant site, plant capacity, and year.  The data set 
will consist of records representing each township in Alberta, with fields
representing township ID, plant location, plant capacity, year, cycle time,
and harvest cost.

@@

All costs and prices used in this study are in Canadian dollars (\$).
At the time of writing
(`r sprintf("%s",fx[1,1])` GMT) , the exchange rate between
Canadian (CAD) and US dollars (USD) was
#`r sprintf("%.4f",fx[1,2])` 
0.8031 USD/CAD.

@@

The bulk of the effort expended on this study was spent in creating
township-level summaries for use by the model.  The model itself is
very simple.  We assume that the residue harvesters follow a kind of
greedy optimization.  They take the cheapest residue first, and stop
when they have enough residue to satisfy annual plant
requirements. For each combination of plant location and year, all the
townships in Alberta were sorted by ascending cycle time (equivalently
tranport cost).  The model works through this list, accumulating
harvest residue recovered from the harvest area associated for the
township and year, and provides a file showing cycle time, total cost,
and cumulative residue recoverd.  Once this file is created, fairly
simple data manipulations are undertaken to produce the results.  See
the pseudocode in section 3.8.1 for a summary.

@@

\newpage 
### Pseudocode

```{r, eval = FALSE}
OBTAIN township-level summary
FOR each plant site in township-level summary
    FOR each year
        SORT township-level summary by ascending cycle time
        SET accumulated biomass to 0
        FOR each capacity
           SELECT records from township-level summaries
               corresponding to plant site, year, and capacity
           REPEAT
               READ township biomass from 
                   sorted township list
               COMPUTE accumulated biomass as
                   current accumulated biomass + township biomass
               WRITE plant site, year, capacity, township id,
                   cycle time, cost, accumulated biomass
            UNTIL accumulated biomass is greater than or equal to
               capacity
        ENDFOR    
    ENDFOR
ENDFOR
CREATE summary graphics and tables

@@

The data compiled to the township-level are currently available on
github(https://github.com/gwa-uab/histressup/tree/main/data).  Before
publication it will be placed on ERA with a DOI.
